---
title: "Tooth Growth Analysis"
author: "Y. Zheng"
date: "Friday, October 14th, 2016"
output:
    pdf_document: default
keep_md: yes
html_document: null
---
    
## Loading Data
```{r}
# Extracting Tooth Growth data and ggplot2 library
library(datasets)
library(ggplot2)
data(ToothGrowth)
```

## Summary of the data
We investigate the relationships between len and each supp as well as len and each dose via exploratory data analysis.

```{r}
# Look at how len varies for each supp
qplot(supp, len, data = ToothGrowth, geom = "boxplot") + 
    labs(title = "Box-plot of len for each supp")


# Look at how len varies for each dose
ToothGrowth$dose = factor(ToothGrowth$dose)
qplot(dose, len, data = ToothGrowth, geom = "boxplot") + 
    labs(title = "Box-plot of len for each dose")
```

From exploratory data analysis, we expect that OJ is the more effective supplement for tooth growth.
We also expect higher dosage to correspond to increased tooth growth (len).

## Hypothesis Testing
We will divide our test into 2 portions: one for supplement and one for dosage.

### Supplement Analysis
Definitions:
    
* ${\mu}_{OJ}$ = population mean of len using OJ supplement
* ${\mu}_{VC}$ = population mean of len using VC supplement

From our exploratory plots, we make the following hypothesis:  
**$H_0: {\mu}_{OJ}-{\mu}_{VC}\leq0$**  
**$H_1: {\mu}_{OJ}>{\mu}_{VC}$**
    
```{r}
t.test(len ~ supp, paired = FALSE, var.equal = FALSE, 
       data = ToothGrowth, alternative = "greater")
```

p-value = 0.03032  
95% confidence interval = [0.4682687, $\infty$]

The 95% confidence interval is entirely positive.  
The p-value perspective tells us that the $\P(Observation|H_0) = 0.03032 < 0.05$.  
*We have sufficient evidence to reject $H_0$ in preference for $H_1: {\mu}_{OJ}>{\mu}_{VC}$*.  

### Dosage Analysis
Definitions:
    
    * ${\mu}_{0.5}$ = population mean of len with 0.5 dose
* ${\mu}_{1}$ = population mean of len with 1 dose
* ${\mu}_{2}$ = population mean of len with 2 dose

Similarly, we expect 1 dose has greater len than 0.5 dose:  
**$H_0: {\mu}_{0.5}-{\mu}_{1}\geq0$**  
**$H_1: {\mu}_{0.5}-{\mu}_{1}<0$**
    
```{r}
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, 
       data = ToothGrowth[ToothGrowth$dose!=2,], alternative = "less")
```

p-value = 6.342e-08

The probability of us observing the data given $H_0$ as prior is vanishingly small. 

*We have sufficient evidence to reject $H_0$ in preference for $H_1: {\mu}_{0.5}-{\mu}_{1}<0$.*
    
From the exploratory plot, we expect 2 dose has greater len than 1 dose:
$H_0: {\mu}_{1}-{\mu}_{2}\geq0$
$H_1: {\mu}_{1}-{\mu}_{2}<0$
    
```{r}
t.test(len ~ dose, paired = FALSE, var.equal = FALSE, 
       data = ToothGrowth[ToothGrowth$dose!=0.5,], alternative = "less")
```

p-value = 9.532e-06

The probability of us observing the data given $H_0$ as prior is vanishingly small.  
*We have sufficient evidence to reject $H_0$ in preference for $H_1: {\mu}_{1}-{\mu}_{2}<0$.* 
    
    
**By the *transitive property of inequality*, we conclude that ${\mu}_{2}>{\mu}_{1}>{\mu}_{0.5}$.**
    
## Conclusions and Assumptions
    
### Assumptions
The above analysis makes the following assumptions:
    
* len is normally distributed
* Each sample is an independent and random observation
* Variables between factors are uncorrelated - this justifies the use of unpaired t-test

We use weighted variances and did not assume equal variance.  
This applies to equal variances without loss of generality (WLOG).

### Conclusions

* **Dosage effectiveness in terms of tooth growth is in this order: 0.5 dose < 1 dose < 2 dose**
* **Supplement OG is more effective than supplement VC in assisting tooth growth**
